The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-10-21 paper/figures.tex
paper: begin model #3 section
paper: add HJ correction vs free param

\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#1.
    This model has fixed stars, fixed planets, twin binaries;
    Note the break in the $x$ axis.
    If the true planet radius is $r_p$, all planets 
    detected in binaries will have apparent radii $r_a = r_p/\sqrt{2}$.
    Only 1 in 8 selected binaries is actually searchable (see 
    Sec.~\ref{sec:model_1}).
    }
    \label{fig:errcases_model_1}
\end{figure}

@@

>>>>

\newpage
\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#1.
    This model has fixed stars, fixed planets, twin binaries;
    Note the break in the $x$ axis.
    If the true planet radius is $r_p$, all planets 
    detected in binaries will have apparent radii $r_a = r_p/\sqrt{2}$.
    Only 1 in 8 selected binaries is actually searchable (see 
    Sec.~\ref{sec:model_1}).
    }
    \label{fig:errcases_model_1}
\end{figure}

@@

\begin{figure}[!tb]
    \centering
    \includegraphics[width=.6\textwidth]{figures/HJ_correction_inputrate_vs_HJratevalues.pdf}
    \caption{
    $\Lambda_i$ is the occurrence rate integrated over all possible phase 
    space for the $i^{\rm th}$ system type. $\Lambda_2/\Lambda_0=1$ 
    corresponds to an equal number of planets per secondary as per single star;
    $\Lambda_2/\Lambda_0=0$ corresponds to secondaries not having any planets.
    While the true HJ occurrence rate (marginalized over stellar multiplicity, 
    Eq.~\ref{eq:occ_rate}) is highly dependent on $\Lambda_2$, 
    the inferred rate hardly depends on whether secondaries have HJs.
    }
    \label{fig:HJ_correction_inputrate_vs_HJratevalues}
\end{figure}

\begin{figure}[!tb]
    \centering
    \includegraphics[width=.6\textwidth]{figures/HJ_correction_inputrate_vs_HJrate_correction_factor.pdf}
    \caption{
    Ratio of inferred to true HJ rates from 
    Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, for Model \#3.
    A plausible value for $\Lambda_2/\Lambda_0$ is $\approx 0.5$, 
    corresponding to an underestimate by a factor of 1.1 (see text).
    }
    \label{fig:HJ_correction_inputrate_vs_HJrate_correction_factor}
\end{figure}

### 2017-10-21 paper/main.tex
paper: add HJ correction vs free param

\shortauthors{LGB, KM, JNW}
\shorttitle{Toy transit surveys}

@@

The aim of this work is to clarify what biases, if any, stellar binarity 
introduces to occurrence rates inferred from transit surveys.
Ignoring binarity leads to diluted planetary radii, overestimated 
detection efficiencies, and an undercounted number of selected stars.
These effects skew occurrence rate measurements in different directions, and 
we develop simplified analytic and numerical models to clarify which are the 
most important.
For models in which all planets have the same radii, we find that 
ignoring binarity leads to an inferred occurrence rate at the true radius two 
to three times smaller than the actual rate (an upper limit to the possible 
error).
%However, the direction and size of the error depend on the assumed 
%planet population's radius distribution.
More realistically, using Howard et al. (2012)'s radius distribution and a 
simple stellar model, we find that ignoring binarity leads to a $\sim10-30\%$ 
underestimate of the number of planets per star over most planetary radii.
In particular, the 30\% underestimate applies to hot Jupiter occurrence rates, 
which suggests that the discrepancy between rates measured by {\it Kepler} and 
RV surveys might be attributable to the {\it Kepler} sample containing more 
binaries.

>>>>

\shortauthors{LGB, KM, JNW}
\shorttitle{Binarity and Occurrence Rates}

@@

The aim of this work is to clarify what biases, if any, stellar binarity 
introduces to occurrence rates inferred from transit surveys.
Ignoring binarity leads to diluted planetary radii, overestimated 
detection efficiencies, and an undercounted number of selected stars.
These effects skew occurrence rate measurements in different directions, and 
we develop simplified analytic and numerical models to clarify which are the 
most important.
For models in which all planets have the same radii, we find that 
ignoring binarity leads to an inferred occurrence rate at the true radius two 
to three times smaller than the actual rate (an upper limit to the possible 
error).
%However, the direction and size of the error depend on the assumed 
%planet population's radius distribution.
More realistically, using Howard et al. (2012)'s radius distribution and a 
simple stellar model, we find that ignoring binarity leads to a $\sim10-30\%$ 
underestimate of the number of planets per star over most planetary radii.
In particular, this underestimate applies to hot Jupiter occurrence rates, 
which suggests that at least part of the discrepancy between rates measured by 
{\it Kepler} and RV surveys might come from the {\it Kepler} sample 
containing more binaries.

### 2017-10-21 paper/model_3.tex
paper: add HJ correction vs free param

In this model, as in the previous one, all single and primary stars have 
identical properties.
Only the secondaries have masses, radii, and luminosities that vary between 
systems: $M\propto R \propto L^{1/\alpha}$.
The radii of planets are assigned independently of any host system
property, and are sampled from an intrinsic radius distribution, which we take 
as
\begin{align}
p_r(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2R_\oplus \\
{\rm constant} & \text{for } r\leq2R_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
Following Howard et al. (2012)'s measurement, we take $\delta = -2.92$.
Our ``nominal model'' remains the same: the binary fraction is 0.44, 
$\alpha=\beta=\gamma=0$.
We take $\Lambda_i$, the occurrence rate integrated over all possible phase 
space for the $i^{\rm th}$ system type, to be equal for singles, primaries, 
and secondaries.

For this model, we forgo analytic development and focus only on numerics 
(Sec.~\ref{sec:numerical_methods}).
The nominal results are shown in Fig.~\ref{fig:errcases_model_3_log}.
For the assumed planetary and stellar distributions, the inferred rate is 
underestimated over all radii.
Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
per star with $r>8R_\oplus$, we can compare the true and inferred hot Jupiter 
occurrence rates.
Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
per thousand stars.
The inferred rate is 6.9 per thousand stars.
This means that the inferred rate underestimates the true rate by a 
multiplicative factor of $\sim 1.3$.
This number is suggestively close to the claimed discrepancy between 
hot Jupiter occurrence rates measured by radial velocity and transit surveys 
(Wright et al. 2012).

However, this result only applies under the assumption that $\Lambda_0 = 
\Lambda_1 = \Lambda_2$.
Given that hot Jupiters are thought to be less common around less-massive 
stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
letting single stars and primaries host planets at the same rate.
We show the HJ multiplicative correction factor as a function of 
$\Lambda_2/\Lambda_0$ in Fig BLAH.

>>>>

In this model, as in the previous one, all single and primary stars have 
identical properties.
Only the secondaries have masses, radii, and luminosities that vary between 
systems: $M\propto R \propto L^{1/\alpha}$.
The radii of planets are assigned independently of any host system
property, and are sampled from an intrinsic radius distribution, which we take 
as
\begin{align}
p_r(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2r_\oplus \\
{\rm constant} & \text{for } r\leq2r_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
Following Howard et al. (2012)'s measurement, we take $\delta = -2.92$.
Our ``nominal model'' remains the same: the binary fraction is 0.44, 
$\alpha=\beta=\gamma=0$.
We take $\Lambda_i$, the occurrence rate integrated over all possible phase 
space for the $i^{\rm th}$ system type, to be equal for singles, primaries, 
and secondaries.

For this model, we forgo analytic development and focus only on numerics 
(Sec.~\ref{sec:numerical_methods}).
The nominal results are shown in Fig.~\ref{fig:errcases_model_3_log}.
For the assumed planetary and stellar distributions, the inferred rate is 
underestimated over all radii.
Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
per star with $r>8r_\oplus$, we can compare the true and inferred hot Jupiter 
occurrence rates.
Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
per thousand stars.
The inferred rate is 6.9 per thousand stars.
This means that the inferred rate underestimates the true rate by a 
multiplicative factor of $\sim 1.3$.
This number is suggestively close to the claimed discrepancy between 
hot Jupiter occurrence rates measured by radial velocity and transit surveys 
(Wright et al. 2012).

However, this result only applies under the assumption that $\Lambda_0 = 
\Lambda_1 = \Lambda_2$.
If hot Jupiters are less common around less-massive stars, it would be more 
sensible to consider $\Lambda_2<\Lambda_0$, while letting single stars and 
primaries host planets at the same rate.
Therefore in Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, we let 
$\Lambda_2$ vary, and show the resulting inferred and true hot Jupiter 
($r>8r_\oplus$) rates.
We see that the inferred rate is nearly independent of $\Lambda_2$~--~this is 
because most ($<1/10$) secondaries are not searchable, and so their 
completeness fraction is much smaller than that of primaries or single stars.
This means far fewer detected hot Jupiters orbit secondaries, and so they 
hardly affect the inferred rate.
In contrast, the true rate is highly dependent on $\Lambda_2$, as one would 
expect from Eq.~\ref{eq:occ_rate}..

This means that the ``correction factor'' from an inferred rate to a true 
rate, in this case for hot Jupiters but also more generally, depends strongly 
on the intrinsic occurrence of the planet type of interest.
This is illustrated visually in 
Fig.~\ref{fig:HJ_correction_inputrate_vs_HJrate_correction_factor}, which 
shows that for $\Lambda_2 = \Lambda_0$, the inferred rate is underestimated by 
a factor of $\sim1.3$.
However if $\Lambda_2 = 0$, the inferred rate is actually an overestimate, 
by a factor of $\sim1.1$.
Though the true ratio $\Lambda_2/\Lambda_0$ is unknown for hot Jupiters, a 
reasonable estimate might be to imagine that only stars with mass 
$>0.7M_\odot$ might be capable of hosting them.
For our nominal mass ratio distribution, this is $48\%$ of secondaries, so a 
reasonable interpretation of 
Fig.~\ref{fig:HJ_correction_inputrate_vs_HJrate_correction_factor} would be 
that an observer ignoring binarity might underestimate HJ occurrence rates by 
a factor of $\sim1.1$.
Assuming that RV surveys contain far fewer binary systems than transit 
surveys, this suggests that the effect likely contributes towards the HJ rate 
discrepancy.
